---
title: "Analysis of Metabolomics Data"
author: "Jing Ma"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r library, include=FALSE}

rm(list = ls())
knitr::opts_chunk$set(fig.pos = 'H')

library(glasso)
library(glmnet)
library(igraph)
library(corpcor)
library(DEGraph)
library(PathNet)
library(KEGGgraph)
library(graphite)
library(KEGGREST)
# library(devtools)
# devtools::install_github("drjingma/netgsa", build_vignettes=T)
# devtools::install_github("JClavel/glassoFast")
library(netgsa)
library(glassoFast)

# Parallel computing
library(parallel)
library(foreach)
library(doParallel)

set.seed(2016)
source("preprocess_lib.r")

```

# Introduction
In this note, we provide details on the metabolomic data preprocessing and enrichment analysis with different methods, including **NetGSA**, **DEGraph**, **PathNet**, and **CAMERA**. 

## R package
Our analysis depends on the following R packages. Additional functions used are also provided.
```{r package, eval=FALSE}
library(glasso)
library(glmnet)
library(igraph)
library(corpcor)
library(DEGraph)
library(PathNet)
library(KEGGgraph)
library(graphite)
library(KEGGREST)
# library(devtools)
# devtools::install_github("drjingma/netgsa", build_vignettes=T)
# devtools::install_github("JClavel/glassoFast")
library(netgsa)
library(glassoFast)

# Parallel computing
library(parallel)
library(foreach)
library(doParallel)

set.seed(2016)
source("preprocess_lib.r")
```



## Data

```{r T1D data}
dat_57 = read.csv('../Datasets/metabolomic_data.csv', check.names = F)

## Create the data list
dataset=list()
dataset$metab_info = data.frame("Number"=seq(3,ncol(dat_57)),"Compound"=as.character(colnames(dat_57)[-c(1,2)]), stringsAsFactors=FALSE)
dataset$metab_info$Number = dataset$metab_info$Number-2
dataset$sample_info= dat_57[,1:2]
colnames(dataset$sample_info) = c("Number",'Label')
dataset$dat=t(apply(dat_57[,-c(1,2)], 2, as.numeric))
dim(dataset$dat)

```

## KEGG metabolic pathways

We searched for the KEGG IDs for each metabolite, which are used to map the pathways.  
```{r KEGG IDs}

dataset$metab_info$CID = NA
dataset$metab_info$HMDB = NA
dataset$metab_info$MID = NA # Metabolite ID
dataset$metab_info$pubchem = NA

dataset$metab_info$CID[1] = 'C07326'
dataset$metab_info$CID[2] = NA; dataset$metab_info$MID[2] = 'ME053703'
dataset$metab_info$CID[3] = NA; dataset$metab_info$HMDB[3] = 'HMDB0031075'
dataset$metab_info$CID[4] = NA; dataset$metab_info$MID[4] = "ME133502"
dataset$metab_info$CID[5] = NA; dataset$metab_info$pubchem[5] = '13120901'   
dataset$metab_info$CID[6] = NA; dataset$metab_info$pubchem[6] = '28115'
dataset$metab_info$CID[7] = 'C00956'
dataset$metab_info$CID[8] = NA; dataset$metab_info$HMDB[8] = 'HMDB0000337'
dataset$metab_info$CID[9] = NA; dataset$metab_info$HMDB[9] = 'HMDB0001987'
dataset$metab_info$CID[10] = 'C03196'

dataset$metab_info$CID[11] = NA; dataset$metab_info$HMDB[11] = 'HMDB0001863'
dataset$metab_info$CID[12] = 'C00233'
dataset$metab_info$CID[13] = 'C05853'
dataset$metab_info$CID[14] = 'C06474'
dataset$metab_info$CID[15] = NA; dataset$metab_info$MID[15] = 'ME053685';
dataset$metab_info$CID[16] = 'C01089'; dataset$metab_info$HMDB[16] = 'HMDB0000357'
dataset$metab_info$CID[17] = 'C00597'; dataset$metab_info$HMDB[17] = 'HMDB0000807';
dataset$metab_info$CID[18] = 'C00156'; dataset$metab_info$HMDB[18] = 'HMDB0000500'
dataset$metab_info$CID[19] = 'C00989'; dataset$metab_info$HMDB[19] = 'HMDB0000710'
dataset$metab_info$CID[20] = NA; dataset$metab_info$HMDB[20] = 'HMDB0000725'
dataset$metab_info$pubchem[21] = '11003628'
dataset$metab_info$pubchem[22] = '1826'
dataset$metab_info$pubchem[23] = '80642'
dataset$metab_info$CID[24] = 'C05659'
dataset$metab_info$CID[25] = 'C00417'
dataset$metab_info$CID[26] = 'C00020'
dataset$metab_info$CID[27] = 'C06104'
dataset$metab_info$CID[28] = 'C00041'
dataset$metab_info$CID[29] = 'C00499'
dataset$metab_info$CID[30] = 'C02350'
dataset$metab_info$CID[31] = 'C00026'
dataset$metab_info$CID[32] = 'C00872'
dataset$metab_info$CID[33] = 'C00259'
dataset$metab_info$CID[34] = 'C00532'
dataset$metab_info$CID[35] = NA; 
dataset$metab_info$CID[36] = 'C00219' #Arachidonic acid
dataset$metab_info$CID[37] = 'C00152'
dataset$metab_info$CID[38] = 'C00049'
dataset$metab_info$CID[39] = 'C08261'
dataset$metab_info$CID[40] = 'C00180'
dataset$metab_info$CID[41] = 'C00099'
dataset$metab_info$CID[42] = 'C01753'
dataset$metab_info$CID[43] = 'C11118'; dataset$metab_info$MID[43] = 'ME133022';
dataset$metab_info$CID[44] = 'C01571'
dataset$metab_info$CID[45] = 'C00187'
dataset$metab_info$CID[46] = 'C00695'
dataset$metab_info$CID[47] = 'C00158'
dataset$metab_info$CID[48] = 'C00327'
dataset$metab_info$CID[49] = NA; dataset$metab_info$pubchem[49] = '9989541'; #Conduritol-beta-epoxide
dataset$metab_info$CID[50] = 'C00791'
dataset$metab_info$CID[51] = 'C00097'
dataset$metab_info$CID[52] = "C01420"
dataset$metab_info$CID[53] = "C00112"
dataset$metab_info$CID[54] = "C00599"; dataset$metab_info$pubchem[54] = "91477"
dataset$metab_info$CID[55]=NA; dataset$metab_info$MID[55] = 'ME053440'; dataset$metab_info$HMDB[55] = "HMDB0000360"
dataset$metab_info$CID[56] = 'C17920';
dataset$metab_info$CID[57] = "C02277"
dataset$metab_info$CID[58] = "C00503"
dataset$metab_info$CID[59] = "C00189"
dataset$metab_info$CID[60] = "C00095"
dataset$metab_info$CID[61] = "C00085"
dataset$metab_info$CID[62] = "NA"
dataset$metab_info$CID[63] = "C01019"
dataset$metab_info$CID[64] = "C00122"
dataset$metab_info$CID[65] = "C01235"; dataset$metab_info$pubchem[65] = '4456'
dataset$metab_info$CID[66] = "C00880"; dataset$metab_info$pubchem[66] = '4136' 
dataset$metab_info$CID[67] = 'C00333'; dataset$metab_info$pubchem[67] = '3627'
dataset$metab_info$CID[68] = NA; dataset$metab_info$HMDB[68] = 'HMDB0029162'; 
dataset$metab_info$CID[69] = "C00198"
dataset$metab_info$CID[70] = "C00031"
dataset$metab_info$CID[71] = "C00103"
dataset$metab_info$CID[72] = "C00092"
dataset$metab_info$CID[73] = "C00191"
dataset$metab_info$CID[74] = "C00025"
dataset$metab_info$CID[75] = "C00064" # L-form
dataset$metab_info$CID[76] = "C00489"
dataset$metab_info$CID[77] = "C00258"
dataset$metab_info$CID[78] = "C00116"
dataset$metab_info$CID[79] = "C05401"; dataset$metab_info$MID[79] = "ME000374"
dataset$metab_info$CID[80] = NA; dataset$metab_info$MID[80] = "EC 3.1.3.21"
dataset$metab_info$CID[81] = "C00037"
dataset$metab_info$CID[82] = "C00160"
dataset$metab_info$CID[83] = NA; dataset$metab_info$pubchem[83] = "4157426"
dataset$metab_info$CID[84] = "C16536"
dataset$metab_info$CID[85] = "C01586"
dataset$metab_info$CID[86] = "C00263"
dataset$metab_info$CID[87] = "C00811"
dataset$metab_info$CID[88] = NA; dataset$metab_info$pubchem[88] = "16639161"
dataset$metab_info$CID[89] = "C00192"
dataset$metab_info$CID[90] = "C00262"
dataset$metab_info$CID[91] = NA; dataset$metab_info$MID[91] = "ME001856";
dataset$metab_info$CID[92] = "C02043"; dataset$metab_info$HMDB[92] = "HMDB0000671" 
dataset$metab_info$CID[93] = "C00294"
dataset$metab_info$CID[94] = "C01711"
dataset$metab_info$CID[95] = NA; dataset$metab_info$MID[95] = 'ME134472'
dataset$metab_info$CID[96] = "C00311"
dataset$metab_info$CID[97] = "C00407"
dataset$metab_info$CID[98] = NA; dataset$metab_info$MID[98] = "ME000375"; dataset$metab_info$pubchem[98] = '151152'
dataset$metab_info$CID[99] = NA; dataset$metab_info$MID[99] = "ME053389"
dataset$metab_info$CID[100] = "C00186" 
dataset$metab_info$MID[101] = "ME000046"; dataset$metab_info$pubchem[101] = "7314"
dataset$metab_info$CID[102] = "C02679"
dataset$metab_info$CID[103] = "C00123" ## L-form
dataset$metab_info$CID[104] = NA; dataset$metab_info$HMDB[104] = "HMDB0000640"
dataset$metab_info$CID[105] = "C01595"
dataset$metab_info$CID[106] = "C06427" # aka Alpha-Linolenic acid
dataset$metab_info$CID[107] = "C00047" # L-lysine
dataset$metab_info$CID[108] = "C07272"
dataset$metab_info$CID[109] = "C00149"
dataset$metab_info$CID[110] = "C00208"
dataset$metab_info$CID[111] = "C00159"
dataset$metab_info$CID[112] = NA; dataset$metab_info$MID[112] = "ME000452";
dataset$metab_info$CID[113] = "C00073"
dataset$metab_info$CID[114] = 'C02989'; dataset$metab_info$HMDB[114] = "HMDB0002005"
dataset$metab_info$CID[115] = NA
dataset$metab_info$CID[116] = "C00137"
dataset$metab_info$CID[117] = "C06424"
dataset$metab_info$CID[118] = "C03137"
dataset$metab_info$CID[119] = NA; dataset$metab_info$MID[119] = "ME054015"; dataset$metab_info$HMDB[119] = "HMDB0000532"
dataset$metab_info$CID[120] = "C00153"
dataset$metab_info$CID[121] = "C00253"
dataset$metab_info$CID[122] = "C02721"
dataset$metab_info$CID[123] = "C00712"
dataset$metab_info$CID[124] = "C00077" #L-form
dataset$metab_info$CID[125] = "C00295"
dataset$metab_info$CID[126] = "C00209"
dataset$metab_info$CID[127] = "C01879" # 5-form, there is also 4-form. 
dataset$metab_info$CID[128] = "C00249"
dataset$metab_info$CID[129] = 'C08362'
dataset$metab_info$CID[130] = 'C00864'
dataset$metab_info$CID[131] = NA; dataset$metab_info$MID[131] = "ME053372"
dataset$metab_info$CID[132] ='C01601'
dataset$metab_info$CID[133] = 'C00079' # L-form
dataset$metab_info$CID[134] = 'C05332'
dataset$metab_info$CID[135] = 'C00346'
dataset$metab_info$CID[136] = NA; dataset$metab_info$MID[136] = 'ME132922'
dataset$metab_info$CID[137] = 'C00408'
dataset$metab_info$CID[138] = 'C00148' # L-form
dataset$metab_info$CID[139] = NA
dataset$metab_info$CID[140] = 'C02067'
dataset$metab_info$CID[141] = 'C00134'
dataset$metab_info$CID[142] = NA
dataset$metab_info$CID[143] = 'C00013'
dataset$metab_info$CID[144] = 'C00022'
dataset$metab_info$CID[145] = 'C00492'
dataset$metab_info$CID[146] = 'C00474'
dataset$metab_info$CID[147] = 'C01685'; dataset$metab_info$HMDB[147] = 'HMDB0000867'
dataset$metab_info$CID[148] = 'C00121' # D-form
dataset$metab_info$CID[149] = 'C00818'; dataset$metab_info$HMDB[149] = 'HMDB0000663' # D-form
dataset$metab_info$CID[150] = 'C00065'
dataset$metab_info$CID[151] = 'C00493'
dataset$metab_info$CID[152] = 'C01984'
dataset$metab_info$CID[153] = 'C08250'
dataset$metab_info$CID[154] = 'C00794'
dataset$metab_info$CID[155] = 'C01530'
dataset$metab_info$CID[156] = 'C05442'
dataset$metab_info$CID[157] = 'C00042'
dataset$metab_info$CID[158] = 'C00089'
dataset$metab_info$CID[159] = 'C00059'
dataset$metab_info$CID[160] = 'C00795' # D-form
dataset$metab_info$CID[161] = 'C06467' # D-form
dataset$metab_info$CID[162] = 'C00245'
dataset$metab_info$CID[163] = 'C16884' # D-form
dataset$metab_info$CID[164] = 'C01620'; dataset$metab_info$HMDB[164] = 'HMDB0000943'; 
dataset$metab_info$CID[165] = 'C00188' # L-form
dataset$metab_info$CID[166] = 'C00178'
dataset$metab_info$CID[167] = "C02477" # alpha-tocopherol
dataset$metab_info$CID[168] = 'C01157'
dataset$metab_info$CID[169] = 'C01083'
dataset$metab_info$CID[170] = NA; dataset$metab_info$MID[170] = 'ME133517'
dataset$metab_info$CID[171] = 'C00078'
dataset$metab_info$CID[172] = 'C00082'
dataset$metab_info$CID[173] = 'C00106'
dataset$metab_info$CID[174] = 'C00086'
dataset$metab_info$CID[175] = 'C00366'
dataset$metab_info$CID[176] = 'C00299'
dataset$metab_info$CID[177] = 'C00183'
dataset$metab_info$CID[178] = 'C00385'
dataset$metab_info$CID[179] = 'C01762'
dataset$metab_info$CID[180] = 'C00379'
dataset$metab_info$CID[181] = 'C00181' # D-form
# print(dataset$metab_info[,c(2:6)])

```

When mapping the pathways, we restricted the pathway size to be at least 5. In addition, we removed the *Metabolic pathways -- Homo sapiens (human)* from enrichment analysis since it is the largest pathway that has 93 members, which cover almost all of the metabolites in this study and is likely to be enriched regardless of the method used. In our analysis, there are in total 65 KEGG metabolic pathways. 

```{r kegg pathway}
kegg_matrix_cpd_pathway = read.csv("../Datasets/KEGG_matrix_cpd_pathway.csv", check.names = FALSE)
cid_shared <- intersect(dataset$metab_info$CID, as.character(kegg_matrix_cpd_pathway[,1]))
dataset <- drop.metabs(dataset, !(dataset$metab_info$CID %in% cid_shared))
dataset$dat <- log2(dataset$dat)
kegg_sub_matrix <- kegg_matrix_cpd_pathway[(kegg_matrix_cpd_pathway[,1] %in% cid_shared),]
rownames(kegg_sub_matrix) =as.character(kegg_sub_matrix[,1])
kegg_sub_matrix = kegg_sub_matrix[,-1]
kegg_sub_matrix = t(as.matrix(kegg_sub_matrix))
kegg_sub_matrix = kegg_sub_matrix[(rowSums(kegg_sub_matrix)>5),]
kegg_sub_matrix = kegg_sub_matrix[-40,] 
dim(kegg_sub_matrix)
kegg_pathways <- kegg_sub_matrix
rownames(dataset$dat) <- dataset$metab_info$CID
kegg_pathways <- kegg_pathways[,match(dataset$metab_info$CID,colnames(kegg_pathways))]

```

KEGG metabolic pathways have a lot of overlap, as shown in the following histogram. 
```{r pathway overlapping, fig.height=7, fig.width=7}
hist(colSums(kegg_sub_matrix), freq = F, xlab='Number of pathways',main='Histogram of pathway overlap')

# ## QQ plot to check whether normality assumption is appropriate for log2 transformed data
# for (j in 1:nrow(dataset$dat)){
#   qqnorm(dataset$dat[j,],main=paste0('Row-',j)); qqline(dataset$dat[j,],col=2);
# }

# save(dataset, kegg_pathways, file='../0_Datasets/metabolomics/T1D_metabolomic_ready.rda')

```

## Network information
Network information is needed for **NetGSA**, **PathNet** and **DEGraph**, which is extracted from KEGG metabolic interactions. Since **PathNet** assess the significance of each pathway independently from the others, we only include binary interactions between metabolites from the same pathway.
Interactions that show up in multiple pathways are recorded multiple times. The network for **DEGraph** is stored in `g`, whereas the network for **NetGSA** is stored in `A`. 

```{r network info}
load('../Datasets/metabolomic_KEGGgraph.rdata')
p <- nrow(dataset$dat)
dataset$metab_info$Number <- as.numeric(gsub('C','',dataset$metab_info$CID))

g <- igraph.from.graphNEL(wholeGraphDir)
g <- as.undirected(g, 'collapse')

rownames(kegg_matrix_cpd_pathway) <- as.character(kegg_matrix_cpd_pathway[,1])
kegg_matrix_cpd_pathway <- kegg_matrix_cpd_pathway[,-1]

el <- get.edgelist(g)
pathways4PathNet <- NULL
for (row in 1:nrow(el)){
  nodes_c <- el[row,]
  tmp <- names(which(colSums(kegg_matrix_cpd_pathway[match(el[row,],rownames(kegg_matrix_cpd_pathway)),])==2))
  if (length(tmp)==0){
    next;
  } else if (length(tmp)>1){
    tmp <- data.frame(t(replicate(length(tmp),el[row,])), 'title'=tmp)
    pathways4PathNet <- rbind(pathways4PathNet,tmp)
  } else {
    cat('Row ',row,' and ',length(tmp),'...\n')
    tmp <- as.data.frame(matrix(c(el[row,],tmp),nrow=1))
    colnames(tmp) <- colnames(pathways4PathNet)
    pathways4PathNet <- rbind(pathways4PathNet,tmp)
  }
}
colnames(pathways4PathNet)[1:2] <- c('id1','id2')
pathways4PathNet$id1 <- as.numeric(gsub('C','',pathways4PathNet$id1))
pathways4PathNet$id2 <- as.numeric(gsub('C','',pathways4PathNet$id2))

dim(pathways4PathNet)

pathways4PathNet <- pathways4PathNet[(pathways4PathNet$title %in% rownames(kegg_pathways)),]
pathways4PathNet$title <- gsub('- Homo sapiens (human)','',pathways4PathNet$title, fixed=TRUE)

g1 <- make_empty_graph(2,directed = FALSE)
V(g1)$name <- setdiff(rownames(dataset$dat), V(g)$name)
g <- igraph::union(g,g1)
g <- induced_subgraph(g, intersect(V(g)$name, dataset$metab_info$CID))

Adj4g <- as.matrix(get.adjacency(g,type="both"))
isSymmetric(Adj4g)
A <- matrix(0,p,p)
rownames(A) <- dataset$metab_info$CID
colnames(A) <- rownames(A)
A[match(rownames(Adj4g), rownames(A)), match(rownames(Adj4g), rownames(A))] <- Adj4g

A4PathNet <- A
rownames(A4PathNet) <- dataset$metab_info$Number
colnames(A4PathNet) <- dataset$metab_info$Number

```

## Enrichment analysis

The following code explains in details how the data are simulated, and enrichment results collected. In brief, we randomly select 20 pathways from which 10\% of metabolites (DC) are set to be *affected* variables. To generate simulated data, we first standardize the original data such that each row (metabolite) has mean zero and unit variance. A non-zero mean signal is then added to each affected metabolite to create *truly* enriched pathways. Lastly, we evaluate the empirical power for each pathway using different methods. All powers are averaged over 200 replications. 

```{r enrichment, eval=FALSE}
nCluster <- 5 # please set the number of clusters for parallel computing
muvals <- 0.5
detect_call <- 0.1
alpha <- 0.05
nreps <- 200
ncond <- 2
p <- nrow(dataset$dat)
npath <- nrow(kegg_pathways)
npath_selected <- 20

# For the selected pathways, assign which variables to affect uniformly.
setList <- lapply(1:nrow(kegg_pathways), function(i) which(kegg_pathways[i,]==1))
names(setList) <- rownames(kegg_pathways)
pathway_sig_id <- sample(1:npath, npath_selected)
isEnrichment <- rep(0, npath)
isEnrichment[pathway_sig_id] <- 1  
genesets <- lapply(1:npath, 
                   function(i) if (i %in% pathway_sig_id){
                     sample(names(setList[[i]]), ceiling(detect_call*length(setList[[i]])))
                     })
variables2affect <- unique(Reduce("c", genesets[pathway_sig_id]))

mu <- vector("list", ncond)
mu[[1]] <- rep(0, p)
mu[[2]] <- muvals * (match(dataset$metab_info$CID, variables2affect, nomatch = 0)>0)

#********************************
#****   Main iterations   *******
#********************************
dat <- vector("list", ncond)
dat[[1]] <- dataset$dat[,(dataset$sample_info$Label=="Diabetic")]    ## p by n1
dat[[2]] <- dataset$dat[,(dataset$sample_info$Label=="Non-Diabetic")]## p by n2
n <- sapply(dat, ncol)
dat <- lapply(dat, function(d) t(scale(t(d))))
dat_combined <- cbind(dat[[1]], dat[[2]])

my.iterations <- function(iters){
  set.seed(iters)
  sigInd <- matrix(0, npath, 12)
  colnames(sigInd) <- c("PathNet-nom", "PathNet-fdr","PathNet-fwer",
                        "camera-nom",  "camera-fdr", "camera-fwer",
                        "DEGraph-nom", "DEGraph-fdr","DEGraph-fwer",
                        "NetGSA-nom", "NetGSA-fdr","NetGSA-fwer")

  # Add the following four lines if shuffling the samples. 
  # ind.1 <- sample(ncol(dataset$dat), n[1])
  # dat <- vector("list", ncond) 
  # dat[[1]] <- dat_combined[,ind.1]
  # dat[[2]] <- dat_combined[,-ind.1]

  for (i in 1:ncond) {
    epsilonVar <- matrix(rnorm(p * n[i]), p, n[i])
    dat[[i]] <- dat[[i]] + matrix(rep(mu[[i]], n[i]), p, n[i]) + 0.5*epsilonVar
  }
  
  ##******************************
  ## (1) *******PathNet*******
  ##******************************
  ## Two-sample t-test to first determine the DE genes;
  pvals4metabs <- sapply(1:p, function(i) t.test(dat[[2]][i,], dat[[1]][i,], var.equal = FALSE)$p.value)
  dat4PathNet <- cbind(dataset$metab_info$Number,-log10(pvals4metabs))
  colnames(dat4PathNet) <- c("CID", "pvals")

  resPathNet <- PathNet(Enrichment_Analysis = TRUE, Contextual_Analysis = FALSE, 
                        DirectEvidence_info = dat4PathNet, Column_DirectEvidence = 2, 
                        Adjacency = A4PathNet, pathway = pathways4PathNet, 
                        n_perm = 100, threshold = alpha, use_sig_pathways  = FALSE)
  
  dict.pathnet <- data.frame("full.name" = sort(rownames(kegg_pathways)), 'short.name' = sort(as.character(resPathNet$enrichment_results$Name)))
  pathnet.tempName <- dict.pathnet$full.name[match(as.character(resPathNet$enrichment_results$Name),as.character(dict.pathnet$short.name))]
  resPathNet$enrichment_results$Name <- as.character(pathnet.tempName)
  resPathNet$enrichment_results <- resPathNet$enrichment_results[match(rownames(kegg_pathways),resPathNet$enrichment_results$Name),]

  sigInd[,1] <- (resPathNet$enrichment_results$p_PathNet<alpha)
  sigInd[,2] <- (p.adjust(resPathNet$enrichment_results$p_PathNet, "BH")<alpha)
  sigInd[,3] <- (p.adjust(resPathNet$enrichment_results$p_PathNet, "holm")<alpha)
  
  ##*******************************  
  ## (2) *******camera*******
  ##*******************************
  classT1D <- 2-as.numeric(dataset$sample_info$Label)
  names(classT1D) <- paste0(dataset$sample_info$Label,dataset$sample_info$Number)
  T1Dprofile <- cbind(dat[[1]],dat[[2]])
  
  design <- cbind(Intercept = 1, Group = classT1D)
  
  resCamera <- limma::camera(T1Dprofile, setList, design) 
  resCamera <- resCamera[match(rownames(kegg_pathways),rownames(resCamera)),]
  sigInd[,4] <- (resCamera$PValue < alpha)
  sigInd[,5] <- (p.adjust(resCamera$PValue, 'BH')<alpha)
  sigInd[,6] <- (p.adjust(resCamera$PValue, 'holm')<alpha)
  
  ##*******************************  
  ## (3) *******DEGraph*******
  ##*******************************
  res.DEGraph <- rep(NA, npath)
  ## Run NetGSA on each pathway
  for (loop_path in 1:npath){
    currentDat <- T1Dprofile[(rownames(T1Dprofile)%in% names(setList[[loop_path]])),]
    g_c <- induced_subgraph(g, names(setList[[loop_path]]))
    X1 <- t(currentDat[, which(dataset$sample_info$Label=='Diabetic')])
    X2 <- t(currentDat[, which(dataset$sample_info$Label=='Non-Diabetic')])
    res.DEGraph[loop_path] <- graph.T2.test(X1, X2, G=igraph.to.graphNEL(g_c))$p.value
  }
  sigInd[,7] <- (res.DEGraph<alpha)
  sigInd[,8] <- (p.adjust(res.DEGraph, "BH")<alpha)
  sigInd[,9] <- (p.adjust(res.DEGraph, "holm")<alpha)
  
  ##*******************************
  ## (4) *******NetGSA*******
  ##*******************************
  ## Estimate the partial correlation matrices
  wAdj <- vector("list", ncond)
  for (k in 1:ncond) {
    lambda_vec = seq(1, 5, 0.1) * sqrt(log(p)/n[k])
    bic_tune <- bic.netEst.undir(x=dat[[k]], one=A, lambda = lambda_vec, rho=0.1 * sqrt(log(p)/n[k]), eta=0.1)
    lambda_opt <- (9+which.min(bic_tune$BIC)) * sqrt(log(p)/n[k]) /10
    fit <- netEst.undir(dat[[k]], one=A, lambda = lambda_opt, rho=0.1 * sqrt(log(p)/n[k]), eta=0.1)
    wAdj[[k]] <- fit$Adj
  }

  out <- NetGSA(A = wAdj, x = T1Dprofile, group = c(rep(1, n[1]), rep(2, n[2])), pathways = kegg_pathways, lklMethod = "REML")
  # Output netgsa results
  sigInd[,10] <- (out$results$pval < alpha)
  sigInd[,11] <- (p.adjust(out$results$pval, "BH") < alpha)
  sigInd[,12] <- (p.adjust(out$results$pval, 'holm')<alpha)

  return(sigInd)
}


cat("In the main loop ... \n")
cl <- makeCluster(nCluster)
registerDoParallel(cl)

res <- foreach(iter = 1:nreps,#.combine='+',
               .packages = c("glasso","glmnet","graphite","igraph","KEGGgraph",
                             "PathNet","limma","DEGraph","org.Hs.eg.db"), .verbose = T) %dopar% {
                               tryCatch({my.iterations(iter)}, error=function(e) NULL)
                             }

stopCluster(cl)

# The average power
sigInd <- Reduce("+", res)/length(res)

```

