---
title: "Analysis of Metabolomics Data"
author: "Jing Ma"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r library, include=FALSE}

rm(list = ls())
knitr::opts_chunk$set(fig.pos = 'H')

library(devtools)
library(graphite)
library(CePa)
library(netgsa)
library(DEGraph)
library(glasso)
library(limma)
library(PathNet)

set.seed(2019)

source("preprocess_lib.r")

```

# Introduction

In this note, we provide details on the metabolomic data preprocessing and enrichment analysis with different methods, including **NetGSA**, **DEGraph**, **PathNet**, **CAMERA** and **CePa**. 

## R package

Our analysis depends on the following R packages and environment. Additional functions needed are included in the file `preprocess_lib.r`. 

```{r package, eval=FALSE}
library(devtools)
library(graphite)
library(CePa)
library(netgsa)
library(DEGraph)
library(limma)
library(PathNet)

set.seed(2019)
source("preprocess_lib.r")
```

```{r session info}
devtools::session_info()
```


## Data

The metabolomic data consist of 71 samples across 181 compounds. These are from 30 diabetic and 41 non-diabetic animals. 

```{r T1D data}
dat_57 = read.csv('../Datasets/metabolomic_ready.csv', check.names = F)

## Create the data list
dataset=list()
dataset$metab_info = data.frame("numeric.id"=seq(3,ncol(dat_57)),"Compound"=as.character(colnames(dat_57)[-c(1,2)]), stringsAsFactors=FALSE)
dataset$metab_info$numeric.id = dataset$metab_info$numeric.id-2
dataset$sample_info= dat_57[,1:2]
colnames(dataset$sample_info) = c("numeric.id",'Label')
dataset$dat=t(apply(dat_57[,-c(1,2)], 2, as.numeric))
dim(dataset$dat)
dataset$dat <- log2(dataset$dat)
table(dataset$sample_info$Label)
```

## KEGG metabolic pathways

We searched for the KEGG IDs for each metabolite, which are used to map the pathways.  
```{r KEGG IDs}
# CID is for KEGG Compound ID
dataset$metab_info$CID = NA
dataset$metab_info$HMDB = NA
dataset$metab_info$MID = NA # Metabolite ID
dataset$metab_info$pubchem = NA

dataset$metab_info$CID[1] = 'C07326'
dataset$metab_info$CID[2] = NA; dataset$metab_info$MID[2] = 'ME053703'
dataset$metab_info$CID[3] = NA; dataset$metab_info$HMDB[3] = 'HMDB0031075'
dataset$metab_info$CID[4] = NA; dataset$metab_info$MID[4] = "ME133502"
dataset$metab_info$CID[5] = NA; dataset$metab_info$pubchem[5] = '13120901'   
dataset$metab_info$CID[6] = NA; dataset$metab_info$pubchem[6] = '28115'
dataset$metab_info$CID[7] = 'C00956'
dataset$metab_info$CID[8] = NA; dataset$metab_info$HMDB[8] = 'HMDB0000337'
dataset$metab_info$CID[9] = NA; dataset$metab_info$HMDB[9] = 'HMDB0001987'
dataset$metab_info$CID[10] = 'C03196'

dataset$metab_info$CID[11] = NA; dataset$metab_info$HMDB[11] = 'HMDB0001863'
dataset$metab_info$CID[12] = 'C00233'
dataset$metab_info$CID[13] = 'C05853'
dataset$metab_info$CID[14] = 'C06474'
dataset$metab_info$CID[15] = NA; dataset$metab_info$MID[15] = 'ME053685';
dataset$metab_info$CID[16] = 'C01089'; dataset$metab_info$HMDB[16] = 'HMDB0000357'
dataset$metab_info$CID[17] = 'C00597'; dataset$metab_info$HMDB[17] = 'HMDB0000807';
dataset$metab_info$CID[18] = 'C00156'; dataset$metab_info$HMDB[18] = 'HMDB0000500'
dataset$metab_info$CID[19] = 'C00989'; dataset$metab_info$HMDB[19] = 'HMDB0000710'
dataset$metab_info$CID[20] = NA; dataset$metab_info$HMDB[20] = 'HMDB0000725'
dataset$metab_info$pubchem[21] = '11003628'
dataset$metab_info$pubchem[22] = '1826'
dataset$metab_info$pubchem[23] = '80642'
dataset$metab_info$CID[24] = 'C05659'
dataset$metab_info$CID[25] = 'C00417'
dataset$metab_info$CID[26] = 'C00020'
dataset$metab_info$CID[27] = 'C06104'
dataset$metab_info$CID[28] = 'C00041'
dataset$metab_info$CID[29] = 'C00499'
dataset$metab_info$CID[30] = 'C02350'
dataset$metab_info$CID[31] = 'C00026'
dataset$metab_info$CID[32] = 'C00872'
dataset$metab_info$CID[33] = 'C00259'
dataset$metab_info$CID[34] = 'C00532'
dataset$metab_info$CID[35] = NA; 
dataset$metab_info$CID[36] = 'C00219' #Arachidonic acid
dataset$metab_info$CID[37] = 'C00152'
dataset$metab_info$CID[38] = 'C00049'
dataset$metab_info$CID[39] = 'C08261'
dataset$metab_info$CID[40] = 'C00180'
dataset$metab_info$CID[41] = 'C00099'
dataset$metab_info$CID[42] = 'C01753'
dataset$metab_info$CID[43] = 'C11118'; dataset$metab_info$MID[43] = 'ME133022';
dataset$metab_info$CID[44] = 'C01571'
dataset$metab_info$CID[45] = 'C00187'
dataset$metab_info$CID[46] = 'C00695'
dataset$metab_info$CID[47] = 'C00158'
dataset$metab_info$CID[48] = 'C00327'
dataset$metab_info$CID[49] = NA; dataset$metab_info$pubchem[49] = '9989541'; #Conduritol-beta-epoxide
dataset$metab_info$CID[50] = 'C00791'
dataset$metab_info$CID[51] = 'C00097'
dataset$metab_info$CID[52] = "C01420"
dataset$metab_info$CID[53] = "C00112"
dataset$metab_info$CID[54] = "C00599"; dataset$metab_info$pubchem[54] = "91477"
dataset$metab_info$CID[55]=NA; dataset$metab_info$MID[55] = 'ME053440'; dataset$metab_info$HMDB[55] = "HMDB0000360"
dataset$metab_info$CID[56] = 'C17920';
dataset$metab_info$CID[57] = "C02277"
dataset$metab_info$CID[58] = "C00503"
dataset$metab_info$CID[59] = "C00189"
dataset$metab_info$CID[60] = "C00095"
dataset$metab_info$CID[61] = "C00085"
dataset$metab_info$CID[62] = "NA"
dataset$metab_info$CID[63] = "C01019"
dataset$metab_info$CID[64] = "C00122"
dataset$metab_info$CID[65] = "C01235"; dataset$metab_info$pubchem[65] = '4456'
dataset$metab_info$CID[66] = "C00880"; dataset$metab_info$pubchem[66] = '4136' 
dataset$metab_info$CID[67] = 'C00333'; dataset$metab_info$pubchem[67] = '3627'
dataset$metab_info$CID[68] = NA; dataset$metab_info$HMDB[68] = 'HMDB0029162'; 
dataset$metab_info$CID[69] = "C00198"
dataset$metab_info$CID[70] = "C00031"
dataset$metab_info$CID[71] = "C00103"
dataset$metab_info$CID[72] = "C00092"
dataset$metab_info$CID[73] = "C00191"
dataset$metab_info$CID[74] = "C00025"
dataset$metab_info$CID[75] = "C00064" # L-form
dataset$metab_info$CID[76] = "C00489"
dataset$metab_info$CID[77] = "C00258"
dataset$metab_info$CID[78] = "C00116"
dataset$metab_info$CID[79] = "C05401"; dataset$metab_info$MID[79] = "ME000374"
dataset$metab_info$CID[80] = NA; dataset$metab_info$MID[80] = "EC 3.1.3.21"
dataset$metab_info$CID[81] = "C00037"
dataset$metab_info$CID[82] = "C00160"
dataset$metab_info$CID[83] = NA; dataset$metab_info$pubchem[83] = "4157426"
dataset$metab_info$CID[84] = "C16536"
dataset$metab_info$CID[85] = "C01586"
dataset$metab_info$CID[86] = "C00263"
dataset$metab_info$CID[87] = "C00811"
dataset$metab_info$CID[88] = NA; dataset$metab_info$pubchem[88] = "16639161"
dataset$metab_info$CID[89] = "C00192"
dataset$metab_info$CID[90] = "C00262"
dataset$metab_info$CID[91] = NA; dataset$metab_info$MID[91] = "ME001856";
dataset$metab_info$CID[92] = "C02043"; dataset$metab_info$HMDB[92] = "HMDB0000671" 
dataset$metab_info$CID[93] = "C00294"
dataset$metab_info$CID[94] = "C01711"
dataset$metab_info$CID[95] = NA; dataset$metab_info$MID[95] = 'ME134472'
dataset$metab_info$CID[96] = "C00311"
dataset$metab_info$CID[97] = "C00407"
dataset$metab_info$CID[98] = NA; dataset$metab_info$MID[98] = "ME000375"; dataset$metab_info$pubchem[98] = '151152'
dataset$metab_info$CID[99] = NA; dataset$metab_info$MID[99] = "ME053389"
dataset$metab_info$CID[100] = "C00186" 
dataset$metab_info$MID[101] = "ME000046"; dataset$metab_info$pubchem[101] = "7314"
dataset$metab_info$CID[102] = "C02679"
dataset$metab_info$CID[103] = "C00123" ## L-form
dataset$metab_info$CID[104] = NA; dataset$metab_info$HMDB[104] = "HMDB0000640"
dataset$metab_info$CID[105] = "C01595"
dataset$metab_info$CID[106] = "C06427" # aka Alpha-Linolenic acid
dataset$metab_info$CID[107] = "C00047" # L-lysine
dataset$metab_info$CID[108] = "C07272"
dataset$metab_info$CID[109] = "C00149"
dataset$metab_info$CID[110] = "C00208"
dataset$metab_info$CID[111] = "C00159"
dataset$metab_info$CID[112] = NA; dataset$metab_info$MID[112] = "ME000452";
dataset$metab_info$CID[113] = "C00073"
dataset$metab_info$CID[114] = 'C02989'; dataset$metab_info$HMDB[114] = "HMDB0002005"
dataset$metab_info$CID[115] = NA
dataset$metab_info$CID[116] = "C00137"
dataset$metab_info$CID[117] = "C06424"
dataset$metab_info$CID[118] = "C03137"
dataset$metab_info$CID[119] = NA; dataset$metab_info$MID[119] = "ME054015"; dataset$metab_info$HMDB[119] = "HMDB0000532"
dataset$metab_info$CID[120] = "C00153"
dataset$metab_info$CID[121] = "C00253"
dataset$metab_info$CID[122] = "C02721"
dataset$metab_info$CID[123] = "C00712"
dataset$metab_info$CID[124] = "C00077" #L-form
dataset$metab_info$CID[125] = "C00295"
dataset$metab_info$CID[126] = "C00209"
dataset$metab_info$CID[127] = "C01879" # 5-form, there is also 4-form. 
dataset$metab_info$CID[128] = "C00249"
dataset$metab_info$CID[129] = 'C08362'
dataset$metab_info$CID[130] = 'C00864'
dataset$metab_info$CID[131] = NA; dataset$metab_info$MID[131] = "ME053372"
dataset$metab_info$CID[132] ='C01601'
dataset$metab_info$CID[133] = 'C00079' # L-form
dataset$metab_info$CID[134] = 'C05332'
dataset$metab_info$CID[135] = 'C00346'
dataset$metab_info$CID[136] = NA; dataset$metab_info$MID[136] = 'ME132922'
dataset$metab_info$CID[137] = 'C00408'
dataset$metab_info$CID[138] = 'C00148' # L-form
dataset$metab_info$CID[139] = NA
dataset$metab_info$CID[140] = 'C02067'
dataset$metab_info$CID[141] = 'C00134'
dataset$metab_info$CID[142] = NA
dataset$metab_info$CID[143] = 'C00013'
dataset$metab_info$CID[144] = 'C00022'
dataset$metab_info$CID[145] = 'C00492'
dataset$metab_info$CID[146] = 'C00474'
dataset$metab_info$CID[147] = 'C01685'; dataset$metab_info$HMDB[147] = 'HMDB0000867'
dataset$metab_info$CID[148] = 'C00121' # D-form
dataset$metab_info$CID[149] = 'C00818'; dataset$metab_info$HMDB[149] = 'HMDB0000663' # D-form
dataset$metab_info$CID[150] = 'C00065'
dataset$metab_info$CID[151] = 'C00493'
dataset$metab_info$CID[152] = 'C01984'
dataset$metab_info$CID[153] = 'C08250'
dataset$metab_info$CID[154] = 'C00794'
dataset$metab_info$CID[155] = 'C01530'
dataset$metab_info$CID[156] = 'C05442'
dataset$metab_info$CID[157] = 'C00042'
dataset$metab_info$CID[158] = 'C00089'
dataset$metab_info$CID[159] = 'C00059'
dataset$metab_info$CID[160] = 'C00795' # D-form
dataset$metab_info$CID[161] = 'C06467' # D-form
dataset$metab_info$CID[162] = 'C00245'
dataset$metab_info$CID[163] = 'C16884' # D-form
dataset$metab_info$CID[164] = 'C01620'; dataset$metab_info$HMDB[164] = 'HMDB0000943'; 
dataset$metab_info$CID[165] = 'C00188' # L-form
dataset$metab_info$CID[166] = 'C00178'
dataset$metab_info$CID[167] = "C02477" # alpha-tocopherol
dataset$metab_info$CID[168] = 'C01157'
dataset$metab_info$CID[169] = 'C01083'
dataset$metab_info$CID[170] = NA; dataset$metab_info$MID[170] = 'ME133517'
dataset$metab_info$CID[171] = 'C00078'
dataset$metab_info$CID[172] = 'C00082'
dataset$metab_info$CID[173] = 'C00106'
dataset$metab_info$CID[174] = 'C00086'
dataset$metab_info$CID[175] = 'C00366'
dataset$metab_info$CID[176] = 'C00299'
dataset$metab_info$CID[177] = 'C00183'
dataset$metab_info$CID[178] = 'C00385'
dataset$metab_info$CID[179] = 'C01762'
dataset$metab_info$CID[180] = 'C00379'
dataset$metab_info$CID[181] = 'C00181' # D-form
# print(dataset$metab_info[,c(2:6)])

```

When mapping the pathways, we restricted the pathway size to be at least 5. In addition, we removed the *Metabolic pathways -- Homo sapiens (human)* from enrichment analysis since it is the largest pathway that has 93 members, which cover almost all of the metabolites in this study and is likely to be enriched regardless of the method used. In our analysis, there are in total 65 KEGG metabolic pathways. 

```{r KEGG pathway}
kegg_matrix_cpd_pathway <- read.csv("../../../../arXiv/BladderCancer/0_Data/KEGG_matrix_cpd_pathway.csv", check.names = FALSE)
kegg_matrix_cpd_pathway[,1] <- as.character(kegg_matrix_cpd_pathway[,1])
kegg_pathways <- lapply(2:ncol(kegg_matrix_cpd_pathway), function(j) kegg_matrix_cpd_pathway[which(kegg_matrix_cpd_pathway[,j]==1),1])
names(kegg_pathways) <- colnames(kegg_matrix_cpd_pathway)[-1]
if (length(grep(pattern = "Metabolic pathways", names(kegg_pathways)))>0 ){
  kegg_pathways <- kegg_pathways[-grep(pattern = "Metabolic pathways", names(kegg_pathways))]
}
# Modify pathway name to make it easier for later. 
names(kegg_pathways) <- gsub('- Homo sapiens (human)','',names(kegg_pathways), fixed=TRUE)

# check for overlapping metabolites between data and pathway
cid_shared <- intersect(dataset$metab_info$CID, unique(unlist(kegg_pathways)))
dataset <- drop.metabs(dataset, !(dataset$metab_info$CID %in% cid_shared))
rownames(dataset$dat) <- dataset$metab_info$CID
dataset$metab_info$numeric.id <- as.numeric(gsub('C','',dataset$metab_info$CID))
kegg_paths_filtered <- lapply(kegg_pathways, function(a) intersect(a, cid_shared))

# Keep pathways with at least 5 members
if (min(sapply(kegg_paths_filtered, length))<5){
  kegg_paths_filtered <- kegg_paths_filtered[which(sapply(kegg_paths_filtered, length)>5)]
}
```

## Network information
Network information for **NetGSA**, **PathNet**, **DEGraph** and **CePa** is extracted from KEGG metabolic interactions using **KEGGgraph**. Note the metabolic network is sparse and several pathways have zero edges. In the case of an empty pathway network, the test in **DEGraph** is equivalent to Hotelling's T-squared test. For **CePa**, pathways with fewer than two edges are removed in enrichment analysis because centrality measures are unreliable. 

```{r network info}
load('../Datasets/metabolomic_KEGGgraph.rdata')

cepa_paths <- prepMetablicPathways(kegg_paths_filtered, "CePa", wholeGraphDir, dataset$metab_info)
netgsa_paths <- prepMetablicPathways(kegg_paths_filtered, "netgsa", wholeGraphDir, dataset$metab_info)
deGraph_paths <- prepMetablicPathways(kegg_paths_filtered, "DEGraph", wholeGraphDir, dataset$metab_info)
camera_paths <- prepMetablicPathways(kegg_paths_filtered, "camera", wholeGraphDir, dataset$metab_info)

PathNet_paths <- prepMetablicPathways(kegg_pathways, "PathNet", wholeGraphDir, dataset$metab_info)
PathNet_paths$pathways <- PathNet_paths$pathways[(PathNet_paths$pathways$title %in% names(kegg_paths_filtered)),]
```

To generate differentially enriched metabolites, we randomly selected 20 out of 65 pathways from which 20\% of metabolites (DC) are set to be *affected* variables. 

```{r affected metabolites, eval=FALSE}
npath <- length(kegg_paths_filtered)
npath_selected <- 20
detect_call <- 0.2

set.seed(2019)
# For the selected pathways, assign which variables to affect uniformly.
pathway_sig_id <- sample(1:npath, npath_selected)
variableSets <- lapply(kegg_paths_filtered[pathway_sig_id], function(a) sample(a, ceiling(detect_call*length(a))))
names(variableSets)  <- NULL
metabs2affect <- unique(unlist(variableSets))

# save(dataset, kegg_paths_filtered, metabs2affect,
#      cepa_paths, netgsa_paths, deGraph_paths, camera_paths, PathNet_paths, 
#      file="../6_Results_New/metabolomics/metabolomics_ready.rda")

```

## Enrichment analysis

The following example code explains in details how the data are simulated, and enrichment results collected. In brief, we first standardize the original data such that each row (metabolite) has mean zero and unit variance. A non-zero mean signal is then added to each affected metabolite to create *truly* enriched pathways. 

Node the example code set mean signal as (`muvals <- 0.5`). In our numerical results, we varied mean signal from 0.1 to 1. Lastly, the following code was run 1000 times by varying `jid` from 1 to 1000. We evaluated the empirical power for each pathway using the proportion of significant discoveries over 1000 replications. 

```{r enrichment, eval=FALSE}
# Load data and write session Info
filename <- "metabolomics_ready.rda"
newFileName <- gsub('.rda','',filename)
newFileName <- gsub('_ready','', newFileName)
load(paste0(OutFolder, filename))

writeLines(capture.output(devtools::session_info()), paste0(newFileName,"_sessionInfo.txt"))

# General parameters
DEBUG <- "p0"
ncond <- 2
muvals <- 0.5
today <- '20190429'

mu <- vector("list", ncond)
mu[[1]] <- rep(0, nrow(dataset$metab_info))
mu[[2]] <- muvals * (match(dataset$metab_info$CID, metabs2affect, nomatch = 0)>0)
npath <- length(kegg_paths_filtered)
path.names <- names(kegg_paths_filtered)

dat <- vector("list", ncond)
dat[[1]] <- dataset$dat[,(dataset$sample_info$Label=="Diabetic")]    ## p by n1
dat[[2]] <- dataset$dat[,(dataset$sample_info$Label=="Non-Diabetic")]## p by n2
dat <- lapply(dat, function(d) t(scale(t(d))))
dat_combined <- cbind(dat[[1]], dat[[2]])
n <- sapply(dat, ncol)
p <- nrow(dataset$metab_info)

#********************************
#****   Main iterations   *******
#********************************

my.iterations <- function(iters){
  set.seed(iters)
  col.names <- c("CAMERA", 
                 "DEGraph", 
                 "NetGSA2", 
                 "NetGSA", 
                 "PathNet", 
                 "cepa.ORA.equal.weight", 
                 "cepa.ORA.in.degree",
                 "cepa.ORA.out.degree",
                 "cepa.ORA.betweenness",
                 "cepa.ORA.in.reach",
                 "cepa.ORA.out.reach",
                 "cepa.GSA.equal.weight", 
                 "cepa.GSA.in.degree",
                 "cepa.GSA.out.degree",
                 "cepa.GSA.betweenness",
                 "cepa.GSA.in.reach",
                 "cepa.GSA.out.reach")
  sigInd <- matrix(NA, npath, length(col.names))
  rownames(sigInd) <- path.names
  colnames(sigInd) <- col.names
  
  # Add the following four lines if shuffling the samples. 
  dat <- vector("list", ncond)
  if (DEBUG=="p1"){
    # Permute the sample labels
    ind.1 <- sample(ncol(dataset$dat), n[1])
    dat[[1]] <- dat_combined[,ind.1]
    dat[[2]] <- dat_combined[,-ind.1]
  } else {
    # Use original sample labels
    dat[[1]] <- dat_combined[,seq(1,n[1])]
    dat[[2]] <- dat_combined[,-seq(1,n[1])]
  }
   
  for (i in 1:ncond) {
    epsilonVar <- matrix(rnorm(p * n[i]), p, n[i])
    dat[[i]] <- dat[[i]] + matrix(rep(mu[[i]], n[i]), p, n[i]) + 0.1*epsilonVar
  }
  
  out.obj <- list()
  out.obj$cepa <- 0
  
  ##*******************************  
  ## camera-----
  ##*******************************
  cat('... Running CAMERA ....\n')
  classT1D <- 2-as.numeric(dataset$sample_info$Label)
  names(classT1D) <- paste0(dataset$sample_info$Label,dataset$sample_info$numeric.id)
  T1Dprofile <- cbind(dat[[1]],dat[[2]])
  design <- cbind(Intercept = 1, Group = classT1D)
  
  resCamera <- limma::camera(T1Dprofile, camera_paths$membership, design) 
  sigInd[,1] <- resCamera$PValue[match(path.names,rownames(resCamera))]
  
  ##*******************************  
  ## DEGraph-----
  ##*******************************
  cat('... Running DEGraph ....\n')
  
  for (loop_path in 1:npath){
    index <- camera_paths$membership[[loop_path]]
    pp <- length(index)
    current_data <- lapply(1:ncond, function(a) dat[[a]][index,])
    sigInd[loop_path,2] <- graph.T2.test(t(current_data[[1]]), t(current_data[[2]]), G=deGraph_paths$pathways[[loop_path]])$p.value
  }

  ##*******************************
  ## NetGSA----
  ##*******************************
  cat('... Running NetGSA ....\n')
  ## Estimate the individual adjacency matrix
  
  wAdj <- vector("list", ncond)
  for (k in 1:ncond) {
    lambda_vec <- seq(1, 5, 0.1) * sqrt(log(p)/n[k])
    bic_tune <- bic.netEst.undir(x=dat[[k]], one=netgsa_paths$adjacency, lambda = lambda_vec, rho=0.1 * sqrt(log(p)/n[k]), eta=0.01)
    lambda_opt <- (9+which.min(bic_tune$BIC)) * sqrt(log(p)/n[k]) /10
    fit <- netEst.undir(dat[[k]], one=netgsa_paths$adjacency, lambda = lambda_opt, rho=0.1 * sqrt(log(p)/n[k]), eta=0.01)
    wAdj[[k]] <- fit$Adj
  }
  
  out <- NetGSA(A = wAdj, x = T1Dprofile, group = c(rep(1, n[1]), rep(2, n[2])), pathways = netgsa_paths$membership, lklMethod = "REHE")
  sigInd[,3] <- out$results$pval[match(path.names, as.character(out$results$pathway))]
  
  # Estimate a single adjacency matrix
  lambda_vec <- seq(1, 5, 0.1) * sqrt(log(p)/sum(n))
  bic_tune <- bic.netEst.undir(x=T1Dprofile, one=netgsa_paths$adjacency, lambda = lambda_vec, rho=0.1 * sqrt(log(p)/sum(n)), eta=0.01)
  lambda_opt <- (9+which.min(bic_tune$BIC)) * sqrt(log(p)/sum(n)) /10
  fit <- netEst.undir(T1Dprofile, one=netgsa_paths$adjacency, lambda = lambda_opt, rho=0.1 * sqrt(log(p)/sum(n)), eta=0.01)
  wAdj <- fit$Adj
  
  out <- NetGSA(A = list(wAdj,wAdj), x = T1Dprofile, group = c(rep(1, n[1]), rep(2, n[2])), pathways = netgsa_paths$membership, lklMethod = "REHE")
  sigInd[,4] <- out$results$pval[match(path.names, as.character(out$results$pathway))]
  
  ##******************************
  ## PathNet-----
  # PathNet pathway edges may need to be updated. Not all pathways are returned with the current edges. 
  ##******************************
  cat('... Running PathNet ....\n')
  
  # Two-sample t-test to first determine the DE metabs;
  pvals4metabs <- sapply(1:p, function(i) t.test(dat[[2]][i,], dat[[1]][i,], var.equal = FALSE)$p.value)
  dat4PathNet <- cbind(dataset$metab_info$numeric.id,-log10(pvals4metabs))
  colnames(dat4PathNet) <- c("CID", "pvals")
  
  resPathNet <- PathNet(Enrichment_Analysis = TRUE, Contextual_Analysis = FALSE,
                  DirectEvidence_info = dat4PathNet, Column_DirectEvidence = 2,
                  Adjacency = PathNet_paths$adjacency, pathway = PathNet_paths$pathways,
                  n_perm = 2000, threshold = 0.05, use_sig_pathways  = FALSE)
  resPathNet <- resPathNet$enrichment_results
  dict.pathnet <- data.frame("full.name" = sort(path.names), 
                             'path.name' = sort(as.character(resPathNet$Name)), stringsAsFactors = FALSE)
  resPathNet$Name <- dict.pathnet$full.name[match(as.character(resPathNet$Name),dict.pathnet$path.name)]
  
  sigInd[,5] <- resPathNet$p_PathNet[match(path.names, resPathNet$Name)]

  ## ******************************
  ## (3) CePa----
  ## ******************************
  cat('... Running CePa ORA...\n')
  logFC.ALL <- rowMeans(dat[[2]]) - rowMeans(dat[[1]])
  names(logFC.ALL) <- dataset$metab_info$CID
  pvals4metabs.fdr <- p.adjust(pvals4metabs, "BH")
  logFC.th <- logFC.ALL[which(pvals4metabs.fdr<0.05)]
  if (length(logFC.th)>0){
    out.obj$cepa <- 1
    res.cepa.ORA <- cepa.all(dif=names(logFC.th), bk=dataset$metab_info$CID, pc = cepa_paths$pathways)
    sigInd[match(names(res.cepa.ORA), path.names),6] <- sapply(res.cepa.ORA, function(a) a$equal.weight$p.value)
    sigInd[match(names(res.cepa.ORA), path.names),7] <- sapply(res.cepa.ORA, function(a) a$in.degree$p.value)
    sigInd[match(names(res.cepa.ORA), path.names),8] <- sapply(res.cepa.ORA, function(a) a$out.degree$p.value)
    sigInd[match(names(res.cepa.ORA), path.names),9] <- sapply(res.cepa.ORA, function(a) a$betweenness$p.value)
    sigInd[match(names(res.cepa.ORA), path.names),10] <- sapply(res.cepa.ORA, function(a) a$in.reach$p.value)
    sigInd[match(names(res.cepa.ORA), path.names),11] <- sapply(res.cepa.ORA, function(a) a$out.reach$p.value)
  }
  
  cat('... Running CePa GSA ... \n')
  classlabels <- list()
  classlabels$label <- c(rep("Pos", n[1]), rep("Neg", n[2]))
  classlabels$treatment <- "Pos"
  classlabels$control <- "Neg"
  
  res.cepa.GSA <- cepa.all(mat = cbind(dat[[1]], dat[[2]]), label = classlabels, pc = cepa_paths$pathways)
  sigInd[match(names(res.cepa.GSA), path.names),12] <- sapply(res.cepa.GSA, function(a) a$equal.weight$p.value)
  sigInd[match(names(res.cepa.GSA), path.names),13] <- sapply(res.cepa.GSA, function(a) a$in.degree$p.value)
  sigInd[match(names(res.cepa.GSA), path.names),14] <- sapply(res.cepa.GSA, function(a) a$out.degree$p.value)
  sigInd[match(names(res.cepa.GSA), path.names),15] <- sapply(res.cepa.GSA, function(a) a$betweenness$p.value)
  sigInd[match(names(res.cepa.GSA), path.names),16] <- sapply(res.cepa.GSA, function(a) a$in.reach$p.value)
  sigInd[match(names(res.cepa.GSA), path.names),17] <- sapply(res.cepa.GSA, function(a) a$out.reach$p.value)
  
  out.obj$sigInd <- sigInd
  
  return(out.obj)
}

res <- tryCatch(my.iterations(jid), error=function(e) NULL)

if (!is.null(res)) {
  cat("... Save Output ... \n")
  save(res, file=paste0(paste0(OutFolder,'results/'), newFileName, '.rda'))
}


```

