---
title: "Analysis of Prostate Cancer Data from TCGA"
author: "Jing Ma"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---


```{r library, include=FALSE}

rm(list = ls())

## Packages for different methods
library(devtools)
library(graphite)
library(org.Hs.eg.db)
library(SPIA)
library(ROntoTools)
library(ToPASeq)
library(CePa)
library(topologyGSA)
library(netgsa)
library(DEGraph)
library(glasso)
library(limma)
library(PathNet)

```

# Introduction

In this note, we provide details on the prostate cancer enrichment analysis with different methods, including **CAMERA**, **CePa**, **DEGraph**, **NetGSA**, **PathNet**, **Pathway-Express**, **PRS**, **SPIA** and **topologyGSA**. 

## R package
The following packages are used throughout for the analysis.
```{r package, eval=FALSE}
## Packages for different methods
library(devtools)
library(graphite)
library(org.Hs.eg.db)
library(SPIA)
library(ROntoTools)
library(ToPASeq)
library(CePa)
library(topologyGSA)
library(netgsa)
library(DEGraph)
library(glasso)
library(limma)
library(PathNet)
```

Computations are done in R 3.5.2. Detailed session information can be retrieved via `session_info()`:

```{r session info}
devtools::session_info()
```

## Data

We first load the data object. 

```{r load data}
load("../Datasets/prostatecancer2015_ready.rda")
ls()
```

The variables in the current working environment are 
  
  * `camera_paths`: a list of pathways defined for use in CAMERA.
  * `cepa_paths`: a list of pathway catelogue information defined for use in CePa.
  * `dataset`: a list with values
    + `gene_info`: a data frame of gene information, including gene symbol and gene Entrez ID. 
    + `sample_info`: a data frame of sample meta data where we used `Status` to define the patient's group (1 for cases and 0 for control).     
    + `dat`: the processed data matrix with rows corresponding to genes and columns to samples.
  * `deGraph_paths`: a list of KEGG pathway graph for use in Pathway-Express.
  * `genes2affect_btw`: genes (in Entrez IDs) that are affected under the betweenness design. 
  * `genes2affect_comm`: genes (in Entrez IDs) that are affected under the community design.
  * `genes2affect_nbhd`: genes (in Entrez IDs) that are affected under the neighborhood design.
  * `netgsa_paths`: a list of pathways defined for use in netgsa.
  * `pathNet_paths`: a list of pathways defined for use in PathNet.
  * `pathSelected`: a list of *graphite* pathways.
  * `pe_paths`: a list of pathways defined for use in Pathway-Express.
  * `topoGSA_paths`: a list of pathways defined for use in topologyGSA. 

One can use the *graphite* package to prepare pathways into a format that can be used in **SPIA**. 

```{r prepare spia, eval=FALSE}
# --Run the following to prepare the pathway information for SPIA--
prepareSPIA(pathSelected, "PCA_KEGG")
```

## Enrichment analysis

The goal is to identify KEGG pathways enriched between normal and diseased patients. In our comparative study, we consider the following two scenarios: (1) use the samples as they are and generate signals at the mean level based on the pathway topology; and (2) shuffle the samples and generate mean level signals. The rationale of (2) is that by repeatedly permuting the samples, we expect no difference between the underlying networks.

The following code was used to run all methods in the final enrichment analysis. Note 

  * Since there are a large number of KEGG pathways and a lot of genes, enrichment analysis is done separately for each pathway.
  * The following code sets mean signal via `muvals=0.1`. In our studies, we let the mean signal `muvals` vary from 0.1, 0.2, 0.3, 0.4 and 0.5 to evaluate the empirical powers of different methods. 


```{r start, eval=FALSE}

source("preprocess_lib.r")

set.seed(2019)

OutFolder <- "../6_Results_New/PCA/"
filename <- "prostatecancer2015_ready.rda"
newFileName <- gsub('.rda','',filename)
newFileName <- gsub('_ready','', newFileName)
load(paste0(OutFolder, filename))

writeLines(capture.output(devtools::session_info()), paste0(newFileName,"_sessionInfo.txt"))

# general parameters
DEBUG <- 'btw' # design for pathway deregulation
today <- '20190403'
ncond <- 2
ncores <- 1
muvals <- 0.1

newFileName <- paste0(newFileName, '_', DEBUG, '_perm_mu', muvals*10, '_ncores', ncores, '_', today, '_', jid)

p <- nrow(dataset$dat)
npath <- length(pathSelected)
path.names <- names(pathSelected)
mu <- vector("list", ncond)
mu[[1]] <- rep(0, p)
mu[[2]] <- muvals * (match(dataset$gene_info$EntrezID, base::eval(as.name(paste0("genes2affect_",DEBUG))), nomatch = 0)>0)

##********************************
##****   Main function   *******
##********************************
dat <- vector("list", ncond)
dat[[1]] <- dataset$dat[,(dataset$sample_info$er_status_by_ihc=="Positive")]## p by n1
dat[[2]] <- dataset$dat[,(dataset$sample_info$er_status_by_ihc=="Negative")]## p by n2
dat <- lapply(dat, function(d) t(scale(t(d))))
n <- sapply(dat, ncol)
dat_combined <- cbind(dat[[1]], dat[[2]])

```

Our main function is `my.iterations`, and simultaneously handles the output from all methods. The results of pathway enrichment are saved in the object `out.obj`.

```{r main function, eval=FALSE}
my.iterations <- function(iters){
  set.seed(iters)
  new_dat <- vector("list", ncond)
  
  ## Permute sample lables
  ind.1 <- sample(ncol(dat_combined), n[1])
  new_dat[[1]] <- dat_combined[,ind.1]
  new_dat[[2]] <- dat_combined[,-ind.1]
  
  ## Or use the original sample labels
  #  new_dat[[1]] <- dat_combined[,seq(1,n[1])]
  #  new_dat[[2]] <- dat_combined[,-seq(1,n[1])]
  dat <- new_dat

  ## output
  col.names <- c("PE.noCut", 
                 "PE.Cut", 
                 "SPIA", 
                 "PRS",
                 "cepa.ORA.equal.weight", 
                 "cepa.ORA.in.degree",
                 "cepa.ORA.out.degree",
                 "cepa.ORA.betweenness",
                 "cepa.ORA.in.reach",
                 "cepa.ORA.out.reach",
                 "cepa.GSA.equal.weight", 
                 "cepa.GSA.in.degree",
                 "cepa.GSA.out.degree",
                 "cepa.GSA.betweenness",
                 "cepa.GSA.in.reach",
                 "cepa.GSA.out.reach",
                 "NetGSA2", 
                 "DEGraph", 
                 "topologyGSA", 
                 "topologyGSA.var", 
                 "CAMERA", 
                 "PathNet",
                 "NetGSA")
  sigInd <- matrix(0, npath, length(col.names))
  rownames(sigInd) <- path.names
  colnames(sigInd) <- col.names
  
  out.obj <- list()
  out.obj$pe <- 0
  out.obj$spia <- 0
  out.obj$prs <- 0
  out.obj$cepa <- 0
  
  cat('... Setting up data ....\n')
  
  for (i in 1:ncond) {
    ## Add some random noise to the data
    epsilonVar <- matrix(rnorm(p * n[i]), p, n[i])
    dat[[i]] <- dat[[i]] + matrix(rep(mu[[i]], n[i]), p, n[i]) + 0.1*epsilonVar
  }
  
  ## ******************************
  ## (1) SPIA, Pathway-Express----
  ## ******************************
  cat('... Running PE ....\n')
  
  ## Two-sample t-test with FDR correction to first determine the DE genes;
  pvals4genes <- sapply(1:p, function(i)  t.test(dat[[2]][i,], dat[[1]][i,], var.equal = FALSE)$p.value)
  
  pvals4genes.fdr <- p.adjust(pvals4genes, "BH")
  logFC.ALL <- rowMeans(dat[[2]]) - rowMeans(dat[[1]])
  names(logFC.ALL) <- dataset$gene_info$EntrezID
  
  res.pe.all <- pe(x = logFC.ALL, graphs = pe_paths, ref = dataset$gene_info$EntrezID,  nboot = 2000, verbose = FALSE)
  res.pe.all <- summary(res.pe.all)
  sigInd[,1] <- res.pe.all$pPert[match(path.names,rownames(res.pe.all))]
 
  logFC.th <- logFC.ALL[which(pvals4genes.fdr<0.05)]
  if (length(logFC.th)>0){
    # Use DE genes
    out.obj$spia <- 1
    out.obj$pe <- 1
    
    res.pe <- pe(x = logFC.th, graphs = pe_paths, ref = dataset$gene_info$EntrezID,  nboot = 2000, verbose = FALSE)
    res.pe <- summary(res.pe)
    sigInd[,2] <- res.pe$pComb[match(path.names,rownames(res.pe))]
    
    cat('... Running SPIA ....\n')
    res.spia <- runSPIA(de=logFC.th, all=dataset$gene_info$EntrezID, "hsa_KEGG")
    sigInd[,3] <- res.spia$pG[match(path.names,res.spia$Name)]
  }
  
  ## ******************************
  ## (2) PRS----
  ## ******************************
  cat('... Running PRS ....\n')
  
  if (length(logFC.th)>0){
    out.obj$prs <- 1
    res.prs <- prs(logFC.th, all=dataset$gene_info$EntrezID, pwys = pathSelected, nperm=2000)
    sigInd[,4] <- res.prs$p.value[match(path.names,rownames(res.prs))]
  }
  
  ## ******************************
  ## (3) CePa----
  ## ******************************
  cat('... Running CePa ORA...\n')
  
  # Note cepa DE gene list is named by gene symbols, and the background genes are also named by gene symbols
  names(logFC.ALL) <- dataset$gene_info$symbol
  logFC.th <- logFC.ALL[which(pvals4genes.fdr<0.05)]
  if (length(logFC.th)>0){
    out.obj$cepa <- 1
    res.cepa.ORA <- cepa.all(dif=names(logFC.th), bk=dataset$gene_info$symbol, pc = cepa_paths$pathways)
    sigInd[,5] <- sapply(res.cepa.ORA, function(a) a$equal.weight$p.value)
    sigInd[,6] <- sapply(res.cepa.ORA, function(a) a$in.degree$p.value)
    sigInd[,7] <- sapply(res.cepa.ORA, function(a) a$out.degree$p.value)
    sigInd[,8] <- sapply(res.cepa.ORA, function(a) a$betweenness$p.value)
    sigInd[,9] <- sapply(res.cepa.ORA, function(a) a$in.reach$p.value)
    sigInd[,10] <- sapply(res.cepa.ORA, function(a) a$out.reach$p.value)
  }
  
  cat('... Running CePa GSA ... \n')
  classlabels <- list()
  classlabels$label <- c(rep("Pos", n[1]), rep("Neg", n[2]))
  classlabels$treatment <- "Pos"
  classlabels$control <- "Neg"
  
  res.cepa.GSA <- cepa.all(mat = cbind(dat[[1]], dat[[2]]), label = classlabels, pc = cepa_paths$pathways)
  sigInd[,11] <- sapply(res.cepa.GSA, function(a) a$equal.weight$p.value)
  sigInd[,12] <- sapply(res.cepa.GSA, function(a) a$in.degree$p.value)
  sigInd[,13] <- sapply(res.cepa.GSA, function(a) a$out.degree$p.value)
  sigInd[,14] <- sapply(res.cepa.GSA, function(a) a$betweenness$p.value)
  sigInd[,15] <- sapply(res.cepa.GSA, function(a) a$in.reach$p.value)
  sigInd[,16] <- sapply(res.cepa.GSA, function(a) a$out.reach$p.value)
  
  ##*******************************  
  ## (4) NetGSA, DEGraph, topologyGSA----
  ##*******************************
  cat('... Running NetGSA, DEGraph and topologyGSA ....\n')
  classlabels <- c(rep(1, n[1]), rep(2, n[2]))
  
  for (loop_path in seq(1,npath)){
    cat("current path ...", loop_path, '/', npath, '...\n')
    index = camera_paths$membership[[loop_path]]
    pp <- length(index)
    current_data <- lapply(1:ncond, function(a) dat[[a]][index,])
    oneMat <- netgsa_paths$adjacency[[loop_path]]
    
    ## Estimate the partial correlation matrices
    wAdj <- vector("list", ncond)
    for (k in 1:ncond) {
      lambda_vec <- seq(1, 5, 0.1) * sqrt(log(pp)/n[k])
      rho_star <- 0.1 * sqrt(log(pp)/n[k])
      bic_tune <- bic.netEst.undir(x=current_data[[k]], one=oneMat,  lambda = lambda_vec, rho=rho_star, eta=0.01*(1+pp/2))
      lambda_opt <- (9+which.min(bic_tune$BIC)) * sqrt(log(pp)/n[k]) /10
      fit <- netEst.undir(current_data[[k]], one=oneMat, lambda = lambda_opt, rho=rho_star, eta = 0.01*(1+pp/2))
      wAdj[[k]] <- fit$Adj
    }

    ## Run NetGSA
    B <- matrix(rep(1,pp), nrow=1)
    colnames(B) <- rownames(oneMat)
    rownames(B) <- path.names[loop_path]
    out <- NetGSA(A = wAdj, x = do.call(cbind, current_data), group=classlabels, pathways = B, lklMethod = "REHE")
    sigInd[loop_path,17] <- out$results$pval
    
    # Also run NetGSA assuming the network is shared between the two conditions
    dat_joint <- do.call(cbind, current_data)
    lambda_vec <- seq(1, 5, 0.1) * sqrt(log(pp)/sum(n))
    rho_star <- 0.1 * sqrt(log(pp)/sum(n))
    bic_tune <- bic.netEst.undir(x=dat_joint, one=oneMat,  lambda = lambda_vec, rho=rho_star, eta=0.01*(1+pp/2))
    lambda_opt <- (9+which.min(bic_tune$BIC)) * sqrt(log(pp)/sum(n)) /10
    fit <- netEst.undir(dat_joint, one=oneMat, lambda = lambda_opt, rho=rho_star, eta = 0.01*(1+pp/2))
    wAdj <- fit$Adj
    out <- NetGSA(A = list(wAdj,wAdj), x = dat_joint, group=classlabels, pathways = B, lklMethod = "REHE")
    sigInd[loop_path,23] <- out$results$pval
    
    sigInd[loop_path,18] <- graph.T2.test(X1=t(current_data[[1]]), X2=t(current_data[[2]]), G=deGraph_paths$pathways[[loop_path]])$p.value

    test <- tryCatch(pathway.mean.test(y1=t(current_data[[1]]), y2=t(current_data[[2]]), dag = topoGSA_paths$pathways[[loop_path]], alpha = 0.05, perm.num = 1000), error=function(e) NULL)
    if (is.null(test)){
      sigInd[loop_path,19] <- NA; sigInd[loop_path,20]  <- NA
    } else {
      sigInd[loop_path,19] <- test$p.value
      sigInd[loop_path,20] <- test$p.value.var
    }
  }
  
  ##*******************************  
  ## (5) camera----
  ##*******************************
  cat('... Running CAMERA ....\n')
  
  classPCA2012 <- c(rep(1, n[1]), rep(0, n[2]))
  names(classPCA2012) <- dataset$sample_info$GSMID
  expPCA2012 <- cbind(dat[[1]], dat[[2]])
  rownames(expPCA2012) <- dataset$gene_info$EntrezID
  design <- cbind(Intercept = 1, Group = classPCA2012)
  
  res.camera <- limma::camera(expPCA2012, camera_paths$membership, design) 
  sigInd[,21] <- res.camera$PValue[match(path.names,rownames(res.camera))]
  
  ## ******************************
  ## (6) PathNet----
  ## ******************************
  cat('... Running PathNet ....\n')
  
  logPval <- -log10(pvals4genes)
  dat4PathNet <- cbind(as.numeric(gsub("ENTREZID:","",dataset$gene_info$EntrezID)),logPval)
  colnames(dat4PathNet) <- c("Gene.ID", "ERneg")
  res.PathNet <- PathNet(Enrichment_Analysis = TRUE, Contextual_Analysis = FALSE, 
                         DirectEvidence_info = dat4PathNet, Column_DirectEvidence = 2, 
                         Adjacency = pathNet_paths$adjacency, pathway = pathNet_paths$pathways, 
                         n_perm = 2000, threshold = 0.05, use_sig_pathways  = FALSE)
  
  res.PathNet <- res.PathNet$enrichment_results
  dict.pathnet <- data.frame("full.name" = sort(path.names), 
                             'path.name' = sort(as.character(res.PathNet$Name)), stringsAsFactors = FALSE)
  res.PathNet$Name <- dict.pathnet$full.name[match(as.character(res.PathNet$Name),dict.pathnet$path.name)]
  sigInd[,22] <- res.PathNet$p_PathNet[match(path.names, res.PathNet$Name)]
  
  out.obj$sigInd <- sigInd
  
  return(out.obj)
}


```

To evaluate empirical power of each method, we run the function `my.iterations` multiple times using the following line, where the variable `jid` ranges from 1 to 200. The empirical power of each pathway is calculated as the proportion of runs which report an adjusted $p$-value that is no greater than 0.05 out of the 200 replications. 

```{r replication, eval=FALSE}

res <- tryCatch(my.iterations(jid), error=function(e) NULL)

if (!is.null(res)) {
  cat("... Save Output ... \n")
  save(res, file=paste0(paste0(OutFolder,'results/mu', muvals*10,'/'), newFileName, '.rda'))
}

cat("... DONE! ... \n")

```




